A Comparative Quality Metric for Fuzzing with Logic State Coverage

### 2024-03-05 paper/main.tex
paper: discussion: quality metric for general fuzzing (revisemake) --> replacing 'completeness of targeted fuzzing'

\pp{Normality of a partial logic state}
%
It is crucial to notice that an abnormal behavior \emph{still follows the
regular control flow up until any exception happens}. In \autoref{f:local}, the
abnormal behavior (Behavior C) follows the same control flow to a normal
behavior (Behavior B) until a use-after-free happens, which makes the behavior
eventually abnormal unlike its normal counterpart. This means that, even though
a behavior is known to be normal, it does not suggest much about the normality
of other behaviors \emph{on course for} such.

@@

\pp{Completeness of targeted fuzzing}
%
The completeness of untargeted fuzzing hinges on the observation that the more
program behavior it checks, the less abnormal program behaviors are expected to
be left. This roughly translates to $\Pr(\mathrm{abnormal} \mid \mathrm{new})$
being equal throughout a fuzzing campaign.
%, whereby $\Pr(\mathrm{new} \cap
%\mathrm{abnormal})$ then majorly depends on how high $\Pr(\mathrm{new})$ is;
%proportional to the number of observed program behaviors per time.
%
However, targeted fuzzing (e.g., directed or regression fuzzing)
\emph{prioritizes} code that likely leads to abnormal behaviors, which may result
in a larger $\Pr(\mathrm{abnormal} \mid \mathrm{new})$ at the beginning that slowly
decreases over time.

This affects the fundamental observation for completeness because targeted
fuzzers may reveal more abnormal behaviors even after observing fewer program
behaviors.
%
One possible remedy is refining interestingness to incorporate a complete set of
factors that affects $\Pr(\mathrm{abnormal})$ (e.g., freshness of code).  A
further investigation on interestingness is required.

>>>>

\pp{Normality of a partial logic state}
%
It is crucial to notice that an abnormal behavior \emph{still follows the
regular control flow up until any exception happens}. In \autoref{f:local}, the
abnormal behavior (Behavior C) follows the same control flow to a normal
behavior (Behavior B) until use-after-free, which makes the behavior eventually
abnormal unlike its normal counterpart. 
%
What it implies is that the executions that do not contribute to the growth of
logic state coverage are not entirely wasteful executions because they are still
possible to reveal abnormal behaviors. However, they are unlikely to do so, let
alone revealing \emph{unknown} ones (\autoref{s:fuzz:unfuzz}).

%This means that, even though a behavior is known to be normal, it does not
%suggest much about the normality of other behaviors \emph{on course for} such.

@@

%\pp{Completeness of targeted fuzzing}
%%
%The completeness of untargeted fuzzing hinges on the observation that the more
%interesting program behavior it checks, the less unknown abnormal program
%behaviors are expected to be left. This roughly translates to $\Pr(\mathrm{UA}
%\mid \mathrm{N})$ being equal throughout a fuzzing campaign.
%%, whereby $\Pr(\mathrm{new} \cap
%%\mathrm{abnormal})$ then majorly depends on how high $\Pr(\mathrm{new})$ is;
%%proportional to the number of observed program behaviors per time.
%%
%However, targeted fuzzing (e.g., directed or regression fuzzing)
%\emph{prioritizes} code that likely leads to abnormal behaviors, which may result
%in a larger $\Pr(\mathrm{UA} \mid \mathrm{N})$ at the beginning that slowly
%decreases over time. 
%
%This affects the fundamental observation for completeness because targeted
%fuzzers may reveal more abnormal behaviors even after observing fewer program
%behaviors.
%%
%One possible remedy is refining interestingness to incorporate a complete set of
%factors that affects $\Pr(\mathrm{UA})$ (e.g., freshness of code).  A
%further investigation on interestingness is required.

% FIXME: revise!
\pp{Quality metric for general fuzzing}
%
The legitimacy of logic state coverage hinges on the current conventional
understanding of interesting program behaviors in \emph{untargeted} fuzzing,
namely the behaviors with less repetition. However, recent findings 
\cite{regression,risk,root-causes-age} suggest that various code characteristics
(e.g., the age of code) make some code likelier to cause bugs,
which makes the program behaviors executing such a portion of code have a higher
expectation to reveal unknown abnormal behaviors (i.e., more interesting).
%
\emph{Targeted} fuzzing \cite{regression,directed,risk} attempts to capitalize
the findings by prioritizing such code, while logic state coverage would
under-estimate their quality by only considering less repetition as more
interesting.

For a quality metric that encompasses general (i.e., both untargeted and
targeted) fuzzing, future research needs to address two issues around
interestingness. First, it should refine the interestingness of program
behaviors. Not only should it systemize the factors that decide interestingness,
but also it should assess which factor affects \emph{more} than others by
\emph{how much}.  Second, it should incorporate such refined interestingness
into the quality metric. If it would follow the philisophy of logic state
coverage (i.e., more uninteresting behavior in fewer logic states),
\emph{grouping} uninteresting branches can be an approach as it makes the
program behaviors with uninteresting branches occupy fewer logic states.
